Multi-domain, multilingual translation with latent multi-task group dropout

### 2021-11-15 2022/acl/acl_latex.tex

The variational distribution $Q$ of $m^d$ is defined independently on a layerwise basis; this means that each layer only involves a subset $\Phi_l^d$ of variational parameters. $Q$ is computed as follows:
\begin{align}
  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} & \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d), k) \nonumber \\
  m_l^d(i) & =  \text{$\mathbb{I}$}(i \in \operatorname{Ind}^d), \nonumber
\end{align}
where $\text{SRS}(\pi,k)$ denotes the process of sampling $k$ times without replacement from the distribution $\pi$, and $\mathbb{I}$ is the indicator function. This modeling choice for the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic} that we use below.  Given our assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be computed as:
\begin{align}
\hspace{-2.em}
D(\dots) &= \mathbb{E}_{m^d \sim Q(m^d | d; \Phi)} \text{log} P(y | m^d, x, d, \theta) \nonumber \\
&= \mathbb{E}_{g^d \sim^{\text{i.i.d}} G(0,1)} \big[ \log P(y |\tilde{m}^d, x, d, \theta,) \big] \nonumber
\end{align}
where the generation process $G(0,1)$ is a product of independent Gumbel distributions, yielding:
\begin{align}
  \hspace{-2.em}
  \forall d,  g^d & =  [g_1^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
  \forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1) \nonumber \\
  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \nonumber \\ 
       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \}\label{eq:top-k} \\
\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
\end{align}
For the second term, the derivation is the following:\fyDone{Check this: we have a sum over layers ?Also: R or -R ?} 
\begin{align}
\hspace{-2.em}
  R    &= \operatorname{KL}(Q(m^d | x, d, \Phi)||P(m^d | x, d; \theta)), \nonumber \\ 
	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big). \label{eq:entropy}
\end{align}
We prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
This inequality shows that an upperbound of $R$ is $\sum_{l=1}^L(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$. During training, we thus maximize a sum over layers of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyDone{Check sum} which performs a regularization over the parameters $\Phi^d$ of the variational distribution.

Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into the objective function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}. However, the operator $\operatorname{Top-k}$, which serves to define $\tilde{m}_l^d$ in Equation~\eqref{eq:top-k}, is not differentiable. Therefore, we approximate this function by the $\operatorname{Soft-Top-K}$ function defined as follows:
\begin{align}
\hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
                    0 \leqslant m_i \leqslant 1 \\
  \forall 0 \leqslant i \leqslant n_d\text{-}1 \label{eq:soft-top-k}\\
        1^{T}.m = k
      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \\ \nonumber
& \text{in which} \nonumber \\
H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i). \nonumber 
\end{align}

@@

\section{Experimental settings \label{sec:experiments}}
\subsection{Baselines}
\subsubsection{Multi-domain, multilingual translation}
In both cases, we examine the performance of three following MT systems.
\begin{itemize}
	\item Transformer
	\item Adapter based Transformer
	\item \system{LaMGD} Tranformer
\end{itemize}
For the sake of brevity, we put the detailed descriptions of these systems in Appendix \ref{appendix:c}.
\subsection{Datasets}
\subsubsection{Multi-domain translation}
We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. For the sake of brevity, we leave most of the details of data in Appendix \ref{appendix:c}.
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks. We evaluate models on widely used multilingual translation datasets.
\begin{itemize}
	\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations of \citet{qi18when} between English and eight related languages.
	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages as in \citet{Wang20balancing}.
\end{itemize}
We clarify the notation of the languages in multilingual experiments as follows
\begin{itemize}
	\item Diverse: bos (bosnia), mar(marathi),
         hin(hindi), mkd(macedonia), ell(greek),
         bulgaria(bul), french(fra), korean(kor) 
	\item Related: azerbajiani (aze), belarusian(bel),
          glacian(glg), slovak (slk), turkisk (tur),
		  russian(rus), portugese(por),czech(ces)
\end{itemize}
The detailed description of the datasets is provided in Appendix \ref{appendix:c}.
\section{Results and ablation studies}
\subsection{Multi-domain translation}
According to Table \ref{tab:mdmt}, \system{LaMGD} Tranformer shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters. Moreover, \system{LaMGD} Tranformer achieves equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.28 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}
\subsubsection{Fuzzy domain separation}
\label{ssec:fuzzy}
In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploiting the proximity between domains. The experiment uses the same data as the previous experiment. However, the domain \domain{law} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}. If the multi-domain system made full use of the proximity between \domain{law$_1$} and \domain{law$_2$}, there should be no significant difference between the performance of the system trained with six original domains and with seven domains replacing \domain{law} by \domain{law$_1$} and \domain{law$_2$}. \citet{Pham21revisiting} showed a large gap between two training settings when using residual adapters. We report below the gap between two settings when using \system{LaMGD} Tranformer.

@@

\subsection{Multilingual translation}
According to Table \ref{tab:multilingual}, \system{LaMGD} Tranformer shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively. Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resourced languages in the list. However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} and Language Adapters in O2M-diverse. 
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.80 &7.30&20.80&21.10&39.70&19.80&22.60&15.20&18.91 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.30&6.80&21.10&22.00&39.70&20.00&23.00&15.20&19.01 \\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.40&16.60&28.50&	27.10&43.70&24.60&30.30&25.60&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.10&15.80&28.40&26.80&43.70&24.50&30.20&25.60&25.64\\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.20&4.0&12.70&22.20&31.80&\SB{34.00}&38.30&8.30&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.20&4.00&\SB{13.30}&21.90&32.20&\SB{34.10}&38.50&8.30&20.31 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.10&3.80&12.60&\SB{22.80}&31.80&33.40&38.10&8.10&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.40&9.70&20.50&31.80&37.50&38.70&39.80&19.00&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.50&9.40&20.00&30.60&37.20&38.20&39.30&19.00&27.03\\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.50}&9.60&\SB{21.50}&32.20&37.70&38.60&40.00&18.90&27.75 \\
    \hline
  \end{tabular}
  \caption{Multilingual translation}
  \label{tab:multilingual}
\end{table*}

>>>>

The variational distribution $Q$ of $m^d$ is defined independently on a layerwise basis; this means that each layer only involves a subset $\Phi_l^d$ of variational parameters. $Q$ is computed as follows:
\begin{align}
  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} & \sim \text{SRS}(\operatorname{softmax}(\Phi_l^d), k) \nonumber \\
  m_l^d(i) & =  \text{$\mathbb{I}$}(i \in \operatorname{Ind}^d), \nonumber
\end{align}
where $\text{SRS}(\pi,k)$ denotes the process of sampling $k$ times without replacement from the distribution $\pi$, and $\mathbb{I}$ is the indicator function. This modeling choice for the latent vector $m_l^d$ is motivated by the Gumbel Top-K trick of \citet{Kool19stochastic} that we use below.  Given our assumptions for the prior distribution and our choice for the variational distribution, the two terms in Eq.~\eqref{eq:lower-bound} can be computed as:
\begin{align}
\hspace{-2.em}
D(\dots) &= \mathbb{E}_{m^d \sim Q(m^d | d; \Phi)} \text{log} P(y | m^d, x, d, \theta) \nonumber \\
&= \mathbb{E}_{g^d \sim^{\text{i.i.d}} G(0,1)} \big[ \log P(y |\tilde{m}^d, x, d, \theta,) \big] \nonumber
\end{align}
where the generation process $G(0,1)$ is a product of independent Gumbel distributions, yielding:
\begin{align}
  \hspace{-2.em}
  \forall d,  g^d & =  [g_1^d, \dots, g_L^d],  \text{ with } g_l^d \in \mathbb{R}^{n_p} \nonumber \\
  \forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1) \nonumber \\
  \operatorname{Ind}^d = \{ i_1, \cdots, i_k \} &= \operatorname{Top-k} \ \{ \ g_l^d(0) + \Phi_l^d(0), \cdots, \nonumber \\ 
       & \quad \quad g_l^d(n_p\text{-}1) + \Phi_l^d(n_p\text{-}1) \ \}\label{eq:top-k} \\
\tilde{m}_l^d(p) &= \mathbb{I}(p \in Ind^d). \nonumber 
\end{align}
For the second term, the derivation is the following:\fyDone{Check this: we have a sum over layers ?Also: R or -R ?} 
\begin{align}
\hspace{-2.em}
  R    &= \operatorname{KL}(Q(m^d | x, d, \Phi)||P(m^d | x, d; \theta)), \nonumber \\ 
	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(m_l^d | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big) \nonumber \\ 
	&= -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1,\cdots,i_k | x, d, \Phi) \big] - \text{log}(\#\Delta^{n_p}_k) \big)\nonumber \\
	&\leqslant  -\sum_{l=1}^L \big(\mathbb{H} \big[ Q(i_1 | \Phi_l^d) \big] - \text{log}(\#\Delta^{n_p}_k) \big). \label{eq:entropy}
\end{align}
We prove inequality~\eqref{eq:entropy} in Appendix~\ref{appendix:b}.
This inequality shows that an upperbound of $R$ is $\sum_{l=1}^L(\text{log}(\#\Delta^{n_p}_k) - \mathbb{H}(\text{softmax}(\Phi_l^d)))$ since $i_1 | \Phi_l^d \sim \operatorname{softmax}(\Phi_l^d)$. During training, we thus maximize a sum over layers of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$\fyDone{Check sum} which performs a regularization over the parameters $\Phi^d$ of the variational distribution.

Thanks to the Gumbel Top-K trick, we can move the parameters $\Phi$ into the objective function and get rid of policy gradients, which have been reported to be very unstable \citep{Diederick14auto}. However, the operator $\operatorname{Top-k}$, which serves to define $\tilde{m}_l^d$ in Equation~\eqref{eq:top-k}, is not differentiable. Therefore, we approximate this function by the $\operatorname{Soft-Top-K}$ function defined as follows:
\begin{align}
\hat{m}_l^d(\tau) &= \displaystyle{\mathop{argmin}_{\substack{
                    0 \leqslant m_i \leqslant 1 \\
  \forall 0 \leqslant i \leqslant n_d\text{-}1 \\
        1^{T}.m = k
      }}} -(g_l^d+\Phi_l^d)^{T} . m - \tau H_b(m) \label{eq:soft-top-k}
\end{align}
in which
\begin{align}
H_b(m) &= - \sum_i m_i \text{log}(m_i) + (1-m_i)\text{log}(1-m_i). \nonumber 
\end{align}

@@

\section{Experimental settings \label{sec:experiments}}
\subsection{Baselines}
\subsubsection{Multi-domain translation}
We describe the choice of our systems in multi-domain experiments as follows
\begin{itemize}
	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048, the multi-head attentions contain with 8 heads.
	\item Adapter based Transformer: The intermediate feeforward dimension is set as 2048 as in \citet{Pham21revisiting}.
	\item \system{LaMGD} Tranformer: There is no change in the architecture. We group 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 groups are selected. The number of parameters of variational distributions are $k \times L \times n_d$ which is tiny compared to the size of Transformer.
\end{itemize}

We set the dropout probability as 0.1. We train the multi-domain Transformer model in 200~K iterations with a batchsize of 12~K tokens. We train \system{LaMGD} Tranformer in 300~K iterations with a batchsize of 12~K tokens since \system{LaMGD} Tranformer has to learn both the sub-networks and the parameters of the underlying Transformer. Finally, we plug Domain Adapters to the multi-domain Transformer model and finetune them with 25~K iterations using the same batchsize as the former.
\subsubsection{Multilingual translation}
We describe the choice of our systems in multilingual experiments as follows
\begin{itemize}
	\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 1024, the multi-head attentions contain with 8 heads as in \citet{Wang20balancing}.
	\item Adapter based Transformer: The intermediate feeforward dimension is set as 128 as in \citet{Gong21adaptive}.
	\item \system{LaMGD} Tranformer: There is no change in the architecture. We group 512 nodes in each layer into 16 groups We group 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 groups are selectedof 32 consecutive nodes. For each domain, only 12 groups are selected. 
\end{itemize}

We set the dropout probability as 0.3. We train the multilingual Transformer model in 40~K iterations with a batchsize of 9600 tokens on 16 GPUs as in \citet{Gong21adaptive}. We train \system{LaMGD} Tranformer in 50~K iterations with the same batchsize. Finally, we finetune the Language Adapters in 5~K iterations.
\subsubsection{Hyperparameters}
\label{ssec:hyperparams}
We choose $n_d = 16$ so that the dropout groups' size is not too small or too large. In our future works, we will do more experiments by changing $n_d$ to examine how this hyperparameters affect the performance of the model. The second hyperparameter in \system{LaMGD} is the number of retained groups in each layer, $k$, which we predetermined as 12 in every experiments. By retaining $12/16$ groups, we share $75 \%$ active groups between two domains or languages in average.

Furthermore, the temperature $\tau$ of our $\operatorname{Soft-Top-K}$ operator will be gradually decreased from $0.5$ to $0.2$ as follows
\begin{align*}
\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \}
\end{align*}
in which $r=0.0001$.

Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments. The choice of the values for our hyperparameters was not optimized and selected by suggestions from previous papers \citep{Jang17categorical,Gong21pay,Gong21adaptive}.
\subsection{Datasets}
\subsubsection{Multi-domain translation}
We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. We describe our datasets for multi-domain translation experiment in Table \ref{tab:Corpora-chap4}
\begin{table*}[h!]
  \centering
  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
    \cline{2-7} 
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
  }
\label{tab:Corpora-chap4}
\end{table*}
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks. We evaluate models on widely used multilingual translation datasets.
\begin{itemize}
	\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations of \citet{qi18when} between English and eight related languages.
	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages as in \citet{Wang20balancing}.
\end{itemize}
We clarify the notation of the languages in multilingual experiments as follows
\begin{itemize}
	\item Diverse: bos (bosnia), mar(marathi),
         hin(hindi), mkd(macedonia), ell(greek),
         bulgaria(bul), french(fra), korean(kor) 
	\item Related: azerbajiani (aze), belarusian(bel),
          glacian(glg), slovak (slk), turkisk (tur),
		  russian(rus), portugese(por),czech(ces)
\end{itemize}
Table \ref{tab:Corpora-multilingual} describes the subset of TED8 that we used in our experiments.
\begin{table*}[h!]
  \centering
  \begin{tabular}{l|cccc} 
  \hline
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
    \hline 
    \multirow{8}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 \\
							 & Belarusian & 4.51k & 248 & 664 \\
							 & Glacian & 10.0k & 682 & 1007 \\
							 & Slovak & 61.5k & 2271 & 2445 \\
							 & Turkish & 182k & 4045 & 5029 \\
                             & Russian & 208k & 4814 & 5483 \\
                             & Portuguese & 185k & 4035 & 4855 \\
                             & Czech & 103k & 3462 & 3831 \\
    \hline
    \multirow{8}{*}{Related} & Bosnian & 5.64k & 474 & 463 \\
                             & Marathi & 9.84k & 767 & 1090 \\
                             & Hindi & 18.79k & 854 & 1243 \\
                             & Macedonian & 25.33k & 640 & 438 \\
                             & Greek & 134k & 3344 & 4433 \\
                             & Bulgarian & 174k & 4082 & 5060 \\
                             & French & 192k & 4320 & 4866 \\
                             & Korean & 205k & 4441 & 5637 \\
  \end{tabular}
  \caption{Data Statistics of TED8 Datasets}
\label{tab:Corpora-multilingual}
\end{table*}
\section{Results and ablation studies}
\subsection{Multi-domain translation}
According to Table \ref{tab:mdmt}, \system{LaMGD} Tranformer shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters. Moreover, \system{LaMGD} Tranformer achieves equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.28 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}
\subsubsection{Fuzzy domain separation}
\label{ssec:fuzzy}
In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploiting the proximity between domains. The experiment uses the same data as the previous experiment. However, the domain \domain{law} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}. If the multi-domain system made full use of the proximity between \domain{law$_1$} and \domain{law$_2$}, there should be no significant difference between the performance of the system trained with six original domains and with seven domains replacing \domain{law} by \domain{law$_1$} and \domain{law$_2$}. \citet{Pham21revisiting} showed a large gap between two training settings when using residual adapters. We report below the gap between two settings when using \system{LaMGD} Tranformer.

@@

\subsection{Multilingual translation}
According to Table \ref{tab:multilingual}, \system{LaMGD} Tranformer shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively. Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resourced languages in the list. However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} and Language Adapters in O2M-diverse. 
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34}&38.3&8.3&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20&30.6&37.2&38.2&39.3&19&27.03\\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40&18.9&27.75 \\
    \hline
  \end{tabular}
  \caption{Multilingual translation}
  \label{tab:multilingual}
\end{table*}
